Cooling Requirements for the Vertical Shear Instability in Protoplanetary Disks

### 2015-05-06 application.tex

>>>>

As an application of our linear models, we estimate where in a
PPD do the thermodynamic conditions allow the VSI to
operate. We first model the cooling time $t_c$ based on realistic
disk properties, which yields a non-uniform dimensionless cooling
time $\beta(r,z;\khat)$. We then compare this to the critical cooling time
$\beta_\mathrm{crit}$ derived above. Finally, we explicitly compute 
linear growth rates numerically, using $\beta(r,z;\khat)$, to
characterize growth times expected in a typical PPD.

### 2015-05-06 numerical.tex

\subsection{Vertically non-isothermal disk model} 
%We briefly consider vertically non-isothermal disks in which there is
%a disk surface at finite height. 
\emph{fix this rambling discussion..}
\citetalias{barker15} have argued that vertically isothermal disks are not
ideal for studing the VSI because $\zmax$ is a free 
parameter in numerical calcluations. On the other hand, vertically 
non-isothermal disks have a disk surface at finite height so there is
a natural choice for $\zmax$. This means that higher order body modes and
surface modes are better defined in a vertically non-isothermal disk.

However, as we have seen in Fig. \ref{bcrit_compare1}, the maximum
growth rate becomes insensitive to $\zmax$ for $\beta\gtrsim
\beta_\mathrm{crit}$. This suggests that the characteristic upper
limit on the thermal timescale for the VSI is not sensitive to
$\zmax$, so that vertically isothermal disks are in fact suitable for
the purpose of characterizing the thermodynamic dependence of the
VSI.

@@

Nevertheless, here we briefly consider a vertically non-isothermal
disk model with $(p,q, h)=(0,-1,0.05)$ and $\Gamma=1.0837$. This
value of $\Gamma$ is chosen so that the standard vertical domain size $\zmax =
5H$ effectively corresponds to the disk surface ($\zmax\simeq
H_s$). The disk temperature decreases rapidly with height such that 
$c_s^2(\zmax)\simeq 10^{-4}c_s^2(0)$.    \emph{note that this would imply midplane
temperatures in excess of $10^6$ K so it's completely un-astrophyscial.}

>>>>

\subsection{Vertically non-isothermal disk model} 
%\emph{fix this rambling discussion..}

% \citetalias{barker15} have argued that vertically isothermal disks are not
% ideal for studing the VSI because $\zmax$ is a free 
% parameter in numerical calcluations. On the other hand, vertically 
% non-isothermal disks have a disk surface at finite height so there is
% a natural choice for $\zmax$. This means that higher order body modes and
% surface modes are better defined in a vertically non-isothermal disk.

% However, as we have seen in Fig. \ref{bcrit_compare1}, the maximum
% growth rate becomes insensitive to $\zmax$ for $\beta\gtrsim
% \beta_\mathrm{crit}$. This suggests that the characteristic upper
% limit on the thermal timescale for the VSI is not sensitive to
% $\zmax$, so that vertically isothermal disks are in fact suitable for
% the purpose of characterizing the thermodynamic dependence of the
% VSI.

@@

Here we briefly consider a vertically non-isothermal
disk model with $(p,q, h)=(0,-1,0.05)$ and $\Gamma=1.0837$. This
value of $\Gamma$ is chosen so that the standard vertical domain size $\zmax =
5H\simeq H_s$ effectively corresponds to the disk surface. The disk
temperature decreases rapidly with height such that
$c_s^2(\zmax)\simeq 10^{-4}c_s^2(0)$. A disk surface at finite height
means that higher order body modes and surface modes are better
defined (cf. in vertically isothermal disk models where the disk
boundary $\zmax$ is a free parameter, see \citetalias{barker15}).

% \emph{note that this would
% imply midplane 
% temperatures in excess of $10^6$ K so it's completely un-astrophyscial.}
